Thesis Title

### 2022-06-01 chapters/chapter_5.tex
Updates from Overleaf

Regarding the quality assessment of the assemblies produced (Figure \ref{fig:chap5_figure4}, Supplementary Table S12), ABySS, BCALM2 and minia are the only single k-mer \ac{dBg} assemblers in the collection and were found to mostly underperform relative to their multiple k-mer \ac{dBg} counterparts, generally resulting in more fragmented assemblies, although there were significant differences in performance across samples. Among multiple k-mer assemblers, VelvetOptimiser frequently produced a very high number of contigs of very small size (over 99\% of the contigs not surpassing the minimum length of 1,000 \ac{bp}) and therefore a low N50 (an average of 29,768 \ac{bp} versus a global average of 84,114 \ac{bp}) (Supplementary tables S5-S10). Additionally, ABySS and VelvetOptimizer produced contigs with a very large number of Ns, with an average of 1,019 and 3,035 uncalled bases per assembly, respectively. MetaHipMer2, although having overall average metrics in the two evenly distributed mock samples (ENN and EMS, Supplementary Tables S5-S6) where it was able to run successfully, it severely underperformed in the real samples (ERR2984773 and ERR2935805, Supplementary Tables S7 and S10). Generally, the performance scores of the assemblers decreased considerably for the real samples in comparison with the simulated ones, either with or without error. High utilisation of the reads in the dataset is observed for most assemblers, with on average at least 90\% of the reads mapping back to the assembly, except for ABySS, BCALM2, MetaHipMer2 and VelvetOptimiser whose values are in the range of 46-79\%. Despite an overall good performance, SPAdes produced the highest number of misassembled contigs, with an average of 98 and a maximum of 572 (sample ERR2935805, Supplementary Table S10), in comparison to the global average of 11 misassembled contigs for all assemblers across all samples.

@@

\begin{itemize}
    \item “--minLength”: Value for minimum contig length, in basepairs. By default, this value is set to 1000 basepairs;
    \item “--mapped\_reads\_threshold”: Value for the minimum percentage of a read aligning to the contig to be considered as mapped. By default, this value is set to 75\%;
    \item “--n\_target”: Target value for the N, NA and NG metrics, ranging from 0 to 100\%. By default, this value is set to 50\%;
    \item “--l\_target”: Target value for the L metric, ranging from 0 to 100\%. By default, this value is set to 90\%;
\end{itemize}

@@

@@

>>>>

Regarding the quality assessment of the assemblies produced (Figure \ref{fig:chap5_figure4}, Supplementary Table S12), ABySS, BCALM2 and minia are the only single k-mer \ac{dBg} assemblers in the collection and were found to mostly underperform relative to their multiple k-mer \ac{dBg} counterparts, generally resulting in more fragmented assemblies, although there were significant differences in performance across samples. Among multiple k-mer assemblers, VelvetOptimiser frequently produced a very high number of contigs of very small size (over 99\% of the contigs not surpassing the minimum length of 1,000 \ac{bp}) and therefore a low N50 (an average of 29,768 \ac{bp} versus a global average of 84,114 \ac{bp}) (Supplementary tables S5-S10). Additionally, ABySS and VelvetOptimizer produced contigs with a very large number of Ns, with an average of 1,019 and 3,035 uncalled bases per assembly, respectively. MetaHipMer2, although having overall average metrics in the two evenly distributed mock samples (ENN and EMS, Supplementary Tables S5-S6) where it was able to run successfully, it severely underperformed in the real samples (ERR2984773 and ERR2935805, Supplementary Tables S7 and S10). Generally, the performance scores of the assemblers decreased considerably for the real samples in comparison with the simulated ones, either with or without error. High utilisation of the reads in the dataset is observed for most assemblers, with on average at least 90\% of the reads mapping back to the assembly, except for ABySS, BCALM2, MetaHipMer2 and VelvetOptimiser whose values are in the range of 46-79\%. Despite an overall good performance, SPAdes produced the highest number of misassembled contigs, with an average of 98 and a maximum of 572 (sample ERR2935805, Supplementary \ac{Pls}0), in comparison to the global average of 11 misassembled contigs for all assemblers across all samples.

@@

\begin{itemize}
    \item \textbf{ “--minLength”:} Value for minimum contig length, in basepairs. By default, this value is set to 1000 basepairs;
    \item \textbf{“--mapped\_reads\_threshold”:} Value for the minimum percentage of a read aligning to the contig to be considered as mapped. By default, this value is set to 75\%;
    \item \textbf{“--n\_target”:} Target value for the N, NA and NG metrics, ranging from 0 to 100\%. By default, this value is set to 50\%;
    \item \textbf{“--l\_target”:} Target value for the L metric, ranging from 0 to 100\%. By default, this value is set to 90\%;
\end{itemize}

@@

\subsubsection{Resource requirements differ greatly}

Regarding computational resources, there is a disparity in usage for the evenly and logarithmically distributed samples (Figure \ref{fig:chap5_sup_figure_2}), with the latter having more resource-intensive requirements possibly due mostly to a higher number of reads. The resource usage also varied greatly by assembler, with multiple k-mer dBG (SPAdes, metaSPAdes, MEGAHIT, SKESA, IDBA-UD, GATBMiniaPipeline and Unicycler) having overall higher resource usage. ABySS performance was inconsistent, having reached a maximum of 1412 CPU hours to produce an assembly (sample ERR2984773), resulting in a run time of 35.52 hours.  MetaHipmer2 was the assembler with the highest memory usage, reaching a maximum of 68.7 GB.

@@

\begin{figure*}[]
\centering
\includegraphics[width=\textwidth]{figures/chapter 5/Supplemental Figure 6.png}
\caption{Assembler performance per reference for the ZymoBIOMICS Microbial Community Standards dataset for sample ERR2984773. The best score for each assembler was selected for 3 LMAS runs. The results for each global assembly metric was normalised, with 1 representing the best result, and 0 the worst. The following assemblers are represented: GATBMiniaPipeline: dark blue, IDBA-UD: light blue, MEGAHIT: dark green, metaSPAdes: light green, SKESA: yellow, SPAdes: orange, Unicycler: red.}
\label{fig:chap5_sup_figure_6}
\end{figure*}

\begin{figure*}[]
\centering
\includegraphics[width=\textwidth]{figures/chapter 5/Supplemental Figure 7.png}
\caption{Genome fragmentation for each reference replicon of the ZimoBIOMICS community standards dataset for the logarithmically distributed samples. Genome fragmentation for the 3 LMAS runs is represented by the number of contigs and breadth of coverage of the reference per assembler for the logarithmically distributed samples: LNN (logarithmically distributed without error model, identified by a circle), LHS (logarithmically distributed with Illumina HiSeq error model, identified by a square) and ERR2935805 (real Illumina HiSeq sample, identified by a diamond). Each assembler is identified with the following colour scheme - dark blue: Unicycler, light blue: SPAdes, dark green: SKESA, light green: metaSPAdes, yellow: MEGAHIT, orange: IDBA-UD, red: GATBMiniaPipeline.}
\label{fig:chap5_sup_figure_7}
\end{figure*}

\begin{landscape}
\input{tables/chapter_5/suptable1}
\end{landscape}

### 2022-06-01 chapters/discussion.tex
Updates from Overleaf

\subsubsection{The need for proper benchmarking}

@@

@@

\subsubsection{The use of mock communities}

>>>>

\subsubsection{The need for better assessment}

@@

\paragraph{Performing proper benchmarking of software} \mbox{}\\

In chapter \ref{ch:paper1} the the suitability of SMg for the microbiological diagnosis, with particular emphasis on the bioinformatic process. In total, 3 different bioinformatics pipelines were evaluated to identify those which could provide the clinical microbiologist with the maximum of relevant and accurate information and differences between them: an open-source Unix-based approach, and two commercial alternatives, Basespace and CLC Genomics Workbench. No approach outperformed the other, and very disparate results were obtained for each one. One of the limitations of this study was that, although several approaches were used, no comprehensive assessment of the tools available for each step of the workflow was performed (quality control, removal of host sequences, taxonomic identification, gene detection and \textit{de novo} assembly). Overall, we found that CLC Genomics Workbench has advantages over the other methods as it does not require previous knowledge of Unix-based tools, it is arguably the most user-friendly, but this conclusion is limited to the set of tools evaluated. Currently, no open-source, user-friendly, one-stop, comprehensive metagenomics toolkit with a visual user interface for shotgun metagenomics analysis is available, but efforts are being made to change this paradigm. Alternatives, such as Anvi'o \citep{eren_anvio_2015} for shotgun metagenomics, and Qiime2 \citep{bolyen_reproducible_2019} for metataxonomics, leverage the individual components into a streamlined manner, but still require  command-line familiarity. Still, each individual component of such software should be individually assessed.

@@

\paragraph{The use of mock communities} \mbox{}\\
